---
title: "get_downsampling_object"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(tidyverse)
options(Seurat.memsafe = TRUE)

```


# Load object with previous clusters
```{r}
Mm_umap_subset <- readRDS("../data_output/surat_objects/Mm_umap_integrated_subset_rep.rds")

```

# Identify cells that belong to high counts clusters
```{r}
Mm_umap_subset@meta.data %>%
  ggplot(aes(nFeature_RNA)) +
  geom_histogram() +
  geom_vline(xintercept = 2500) + 
    geom_vline(xintercept = 3500) 

```
```{r}
Mm_umap_subset$is_high <- Mm_umap_subset$nFeature_RNA > 2500
```

Which clusters are high?
```{r}
cluster_by_ishigh <- table(Mm_umap_subset$seurat_clusters, Mm_umap_subset$is_high)

high_clusters <- rownames(cluster_by_ishigh)[apply(cluster_by_ishigh,1,function(x) x[1]<x[2])]

Mm_umap_subset$cluster_is_high <- Mm_umap_subset$seurat_clusters %in% high_clusters
```

Get cells in high clusters
```{r}
cells_in_high_clusters <- Cells(Mm_umap_subset)[Mm_umap_subset$cluster_is_high]
```

Clean memory
```{r}
rm(Mm_umap_subset)
gc()
```

```{r}
get_high_cells_from_seurat <- function(seurat_obj, what_is_high = 2500) {
  seurat_obj$is_high <- seurat_obj$nFeature_RNA > what_is_high
  cluster_by_ishigh <- table(seurat_obj$seurat_clusters, seurat_obj$is_high)
  
  high_clusters <- rownames(cluster_by_ishigh)[apply(cluster_by_ishigh,1,function(x) x[1]<x[2])]
  
  seurat_obj$cluster_is_high <- seurat_obj$seurat_clusters %in% high_clusters
  cells_in_high_clusters <- Cells(seurat_obj)[seurat_obj$cluster_is_high]
  return(cells_in_high_clusters)
}
```


# Subsample these cells in o.g. objects
```{r}
Mm.merged.subset <- readRDS( file = "../data_output/surat_objects/Mm_merged_mito_subset.rds")

#Mm.merged.subset_mito <- saveRDS("../data_output/surat_objects/Mm_merged_mito_subset_mito.rds")

```

```{r}
cells_in_high_clusters_filter <- Cells(Mm.merged.subset) %in% cells_in_high_clusters
```

```{r}
# Get non zero element
Mm.merged.subset.downsampled <- Mm.merged.subset
Mm.merged.subset.downsampled@assays$RNA@counts[,cells_in_high_clusters_filter] <- floor(Mm.merged.subset.downsampled@assays$RNA@counts[,cells_in_high_clusters_filter] * 0.25)

```

```{r}
Mm.merged.subset.downsampled@meta.data$nFeature_RNA_downsampled <- apply(Mm.merged.subset.downsampled@assays$RNA@counts,2,Matrix::nnzero)
Mm.merged.subset.downsampled@meta.data$nCountRNA_downsampled <- apply(Mm.merged.subset.downsampled@assays$RNA@counts,2,sum)


Mm.merged.subset.downsampled@meta.data %>%
  ggplot(aes(nFeature_RNA_downsampled)) +
  geom_histogram() +
  facet_grid(rows = cells_in_high_clusters_filter)


Mm.merged.subset.downsampled@meta.data %>%
  ggplot(aes(nCountRNA_downsampled)) +
  geom_histogram() +
  facet_grid(rows = cells_in_high_clusters_filter)

```
```{r}
Mm.merged.subset.downsampled@meta.data %>%
  ggplot(aes(nFeature_RNA)) +
  geom_histogram() +
  geom_vline(xintercept = 2500) + 
    geom_vline(xintercept = 3500) 

```

```{r}
Mm.merged.subset.downsampled
```


# Downsample all four seurat mito objects 

```{r}
downscale_seurat <- function(seurat_obj, which_cells, how_much = 0.25) {
  # Downscale
  seurat_obj@assays$RNA@counts[,which_cells] <- floor(seurat_obj@assays$RNA@counts[,which_cells] * how_much)
  # Update features
  seurat_obj@meta.data$nFeature_RNA <- apply(seurat_obj@assays$RNA@counts,2,Matrix::nnzero)
  seurat_obj@meta.data$nCount_RNA <- apply(seurat_obj@assays$RNA@counts,2,sum)
  
  return(seurat_obj)
}

get_high_cells_from_seurat <- function(seurat_obj, what_is_high = 2500) {
  seurat_obj$is_high <- seurat_obj$nFeature_RNA > what_is_high
  cluster_by_ishigh <- table(seurat_obj$seurat_clusters, seurat_obj$is_high)
  
  high_clusters <- rownames(cluster_by_ishigh)[apply(cluster_by_ishigh,1,function(x) x[1]<x[2])]
  
  seurat_obj$cluster_is_high <- seurat_obj$seurat_clusters %in% high_clusters
  cells_in_high_clusters <- Cells(seurat_obj)[seurat_obj$cluster_is_high]
  return(cells_in_high_clusters)
}
downsample_seurat_from_file <- function(file_clusters,file_raw) {
  which_cells <- get_high_cells_from_seurat(seurat_obj = readRDS(file_clusters))
  seurat_obj <- downscale_seurat(seurat_obj = readRDS( file = file_raw),which_cells = which_cells)
  
  saveRDS(seurat_obj,file =  paste0(tools::file_path_sans_ext(file_raw),"_downsample.rds")
)
}

```

```{r}
seurat_obj_file_list <- list.files("../data_output/surat_objects", pattern = "mito",full.names = T)
seurat_obj_file_list <- seurat_obj_file_list[!str_detect(seurat_obj_file_list,"rep")]

seurat_umap_file_list <- list.files("../data_output/surat_objects/run_8nov21", pattern = "integrated.rds",full.names = T)
seurat_umap_file_list <- seurat_umap_file_list[c(2,4,3,1)]

 
walk2(seurat_umap_file_list, seurat_obj_file_list, downsample_seurat_from_file)


```

Re-filter objects
```{r}
seurat_obj_file_list <- list.files("../data_output/surat_objects", pattern = "downsample",full.names = T)
seurat_obj_list<- map(seurat_obj_file_list, readRDS)
```

Re-calculate mito content
```{r}

matrix_dir = "../data_raw/scmm_project/G_A1/outs/filtered_feature_bc_matrix/"
features.path <- paste0(matrix_dir, "features.tsv.gz")
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
mito_features <- feature.names$V1[stringr::str_detect(feature.names$V2,"^mt-")]

seurat_obj_list[[2]][["percent.mt"]] <- PercentageFeatureSet(seurat_obj_list[[2]],features = mito_features)
seurat_obj_list[[4]][["percent.mt"]] <- PercentageFeatureSet(seurat_obj_list[[4]],features = mito_features)


```

Re- filter
```{r}
seurat_obj_list <- map(seurat_obj_list, ~ subset(.x, subset = nFeature_RNA > 500 & nFeature_RNA < 7500 & nCount_RNA < 50000 & percent.mt < 5))


```

Save filtered objects
```{r}
walk2(seurat_obj_list, seurat_obj_file_list, saveRDS)

```

# Get separate cluster annotation for high counts clusters
```{r}
get_high_cells_seurat_from_file <- function(file_clusters,file_raw) {
  which_cells <- get_high_cells_from_seurat(seurat_obj = readRDS(file_clusters))

  return(which_cells)
}
get_high_clusters_seurat_from_file <- function(file_clusters,file_raw, what_is_high = 2500) {
  seurat_obj <- readRDS(file_clusters)
  
  seurat_obj$is_high <- seurat_obj$nFeature_RNA > what_is_high
  cluster_by_ishigh <- table(seurat_obj$seurat_clusters, seurat_obj$is_high)
  
  high_clusters <- rownames(cluster_by_ishigh)[apply(cluster_by_ishigh,1,function(x) x[1]<x[2])]
  
  return(high_clusters)
}


seurat_umap_file_list <- list.files("../data_output/surat_objects/run_8nov21", pattern = "integrated.rds",full.names = T)


high_counts_list <- lapply(seurat_umap_file_list, get_high_cells_seurat_from_file)
high_clusters_list <- lapply(seurat_umap_file_list, get_high_clusters_seurat_from_file)

names(high_counts_list) <- basename(seurat_umap_file_list)
names(high_clusters_list) <- basename(seurat_umap_file_list)
saveRDS(high_counts_list, "../data_output/surat_objects/high_counts_list_11nov21.rds")
saveRDS(high_clusters_list, "../data_output/surat_objects/high_clusters_list_11nov21.rds")

```


