---
title: "RGC subclustering 4 replicates DE"
output: html_document
date: '2023-09-07'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(tidyverse)
options(Seurat.memsafe = TRUE)
```
# Radial glia analysis

The aim of this analysis is to sub-cluster (both fzd8+ and fzd8-) radial glial cells  to get a higher resolution clustering of this cell population. Using both samples to get DE genes.

# Preprocessing



Read seurat object
```{r eval = FALSE}
seurat_obj <- readRDS("../data_output/surat_objects/run_8nov21/Mm_merged_mito_mito_integrated.rds")

```
New cluster assignments
```{r eval = FALSE}
DefaultAssay(seurat_obj) <- "integrated"
seurat_obj <- FindNeighbors(object = seurat_obj, dims = 1:20)
seurat_obj <- FindClusters(object = seurat_obj, resolution = 1.1)

```

```{r eval = FALSE}
library(clusterProfiler)
library(org.Mm.eg.db)
DefaultAssay(seurat_obj) <- "RNA"


fzd8 <- bitr("Fzd8",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

seurat_obj <- NormalizeData(seurat_obj)

FeaturePlot(object = seurat_obj,features = fzd8) +
  theme(aspect.ratio = 1) 

VlnPlot(object = seurat_obj, features = fzd8,sort = T) +
  theme(aspect.ratio = 0.75)

DimPlot(object = seurat_obj,label = T) +
  theme(aspect.ratio = 1) 
```


```{r eval = FALSE}
fzd8_clusters <- c("1","6","12","14","3","9")
```

# Filtering

Keep cells from clusters annotated as tbr2+ intermediate progenitor neurons.
```{r eval = FALSE}
seurat_obj <- seurat_obj[,seurat_obj$seurat_clusters %in% fzd8_clusters]
```


# Clustering
```{r eval = FALSE}
DefaultAssay(seurat_obj) <- "integrated"

seurat_obj <- RunPCA(seurat_obj, verbose = FALSE, npcs = 100)

seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:20)

seurat_obj <- FindNeighbors(object = seurat_obj, dims = 1:20)
seurat_obj <- FindClusters(object = seurat_obj, resolution = 1.1)

```
```{r eval = FALSE}
saveRDS(seurat_obj,file = "../data_output/surat_objects/rgc_both_litters_seurat_obj.Rds")
```


```{r}
seurat_obj <- readRDS(file = "../data_output/surat_objects/rgc_both_litters_seurat_obj.Rds")

```


# Visualizing UMAP
```{r}
pdf("../data_output/20230907_rgc_both_litters_umap_fzd8.pdf")
DimPlot(seurat_obj, label = TRUE) + theme(aspect.ratio = 1)

DimPlot(seurat_obj, label = TRUE, split.by = "geno") + theme(aspect.ratio = 1)
dev.off()
```


```{r}
fzd8 <- bitr("Fzd8",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

DefaultAssay(seurat_obj) <- "RNA"
FeaturePlot(seurat_obj,features = fzd8) + theme(aspect.ratio = 1)

```



# Get markers
```{r eval = FALSE}
DefaultAssay(seurat_obj) <- "RNA"
seurat_obj <- NormalizeData(seurat_obj)

seurat_obj_conserved_markers_list <- lapply(levels(Idents(seurat_obj)), FindConservedMarkers, object = seurat_obj,grouping.var = "geno", verbose = FALSE )

names(seurat_obj_conserved_markers_list) <- levels(Idents(seurat_obj))

seurat_obj_conserved_markers_list_df <- do.call(rbind,imap(seurat_obj_conserved_markers_list, function(.x,.y) {
  .x <- rownames_to_column(.x,var = "ensembl")
  .x$cluster <- .y
  return(.x)
}))
saveRDS(seurat_obj_conserved_markers_list_df, "../data_output/surat_objects/rgc_both_litters_conserved_markers_list_df.rds")

top30_markers <- seurat_obj_conserved_markers_list_df %>%
  group_by(cluster) %>%
  top_n(30,XP_avg_log2FC)

library(clusterProfiler)
library(org.Mm.eg.db)

symbol_map <- bitr(top30_markers$ensembl,fromType = "ENSEMBL",toType = "SYMBOL",OrgDb = org.Mm.eg.db)

symbol_map <- symbol_map[!duplicated(symbol_map$ENSEMBL),]

top30_markers <- left_join(top30_markers,symbol_map, by =c("ensembl" = "ENSEMBL"))

write.csv(top30_markers, "../data_output/surat_objects/rgc_all_samples_top30lfc_markers.csv")

```

# Tbr2
```{r}
library(clusterProfiler)
library(org.Mm.eg.db)

tbr2 <- bitr("Eomes",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

FeaturePlot(object = seurat_obj,features = tbr2) +
  theme(aspect.ratio = 1) 

VlnPlot(object = seurat_obj, features = tbr2) +
  theme(aspect.ratio = 0.75)
```


# Fzd8 
```{r}
pdf("../data_output/20230907_rgc_both_litters_umap_fzd8.pdf")

fzd8 <- bitr("Fzd8",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

FeaturePlot(object = seurat_obj,features = fzd8) +
  theme(aspect.ratio = 1)

VlnPlot(object = seurat_obj, features = fzd8) +
  theme(aspect.ratio = 0.75)

FeaturePlot(object = seurat_obj,features = fzd8 , split.by = "geno") +
  theme(aspect.ratio = 1)

VlnPlot(object = seurat_obj, features = fzd8, split.by = "geno") +
  theme(aspect.ratio = 0.75)

dev.off()
```
# Malat
```{r}

malat1 <- bitr("Malat1",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

seurat_obj$malat_percent <- PercentageFeatureSet(seurat_obj,features = malat1)

FeaturePlot(object = seurat_obj,features = "malat_percent") +
  theme(aspect.ratio = 1)

VlnPlot(object = seurat_obj, features = "malat_percent") +
  theme(aspect.ratio = 1)
```

```{r}


```

```{r}
tabled_n_by_clust <- table(seurat_obj@meta.data$seurat_clusters,seurat_obj@meta.data$geno)
relative_tabled_n_by_clust <- tabled_n_by_clust
relative_tabled_n_by_clust[,"WT"] <- relative_tabled_n_by_clust[,"WT"]/sum(relative_tabled_n_by_clust[,"WT"])
relative_tabled_n_by_clust[,"XP"] <- relative_tabled_n_by_clust[,"XP"]/sum(relative_tabled_n_by_clust[,"XP"])


```

```{r}
pdf("../data_output/20230907_rgc_both_litters_abundance.pdf")
seurat_obj@meta.data %>%
  as_data_frame() %>%
  ggplot(aes(seurat_clusters, fill =  geno)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(type = "qual",palette = 7) +
  theme_minimal()

as.data.frame(relative_tabled_n_by_clust) %>%
  mutate(seurat_clusters = Var1,
         geno = Var2) %>%
  ggplot(aes(seurat_clusters,Freq, fill =  geno)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(type = "qual",palette = 7) +
  theme_minimal()
dev.off()
```
```{r}
as.data.frame(relative_tabled_n_by_clust) %>%
  mutate(seurat_clusters = Var1,
         geno = Var2) %>%
  select("Freq","seurat_clusters","geno") %>%
  write_csv(., file = "../data_output/20231011_rgc_both_litters_abundance_table.csv")
```



EdgeR
```{r}
library(scater)
library(edgeR)
sce_obj <- as.SingleCellExperiment(seurat_obj)

abundances <- table(seurat_obj@meta.data$seurat_clusters,seurat_obj@meta.data$sample)

abundances <- unclass(abundances) 


extra_info <- colData(sce_obj)[match(colnames(abundances), sce_obj$sample),]

extra_info$geno <- factor(extra_info$geno)
y_ab <- DGEList(abundances, samples=extra_info)

design <- model.matrix(~ geno, y_ab$samples)

y_ab <- calcNormFactors(y_ab)

y_ab <- estimateDisp(y_ab, design, trend="none")
fit_ab <- glmQLFit(y_ab, design, robust=TRUE, abundance.trend=FALSE)

res <- glmQLFTest(fit_ab, coef=ncol(design))
summary(decideTests(res))

res$table
```


```{r}

as.data.frame(relative_tabled_n_by_clust) %>%
  mutate(seurat_clusters = Var1,
         geno = Var2) %>%
  filter(seurat_clusters %in% rownames(res$table[res$table$logFC > 0.5,])) %>%
  ggplot(aes(seurat_clusters,Freq, fill =  geno)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(type = "qual",palette = 7) +
  theme_minimal() +
  theme(aspect.ratio = 1)

```

# Cell cycle
```{r eval = FALSE}
set.seed(100)

library(scran)
mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", 
    package="scran"))

# Using Ensembl IDs to match up with the annotation in 'mm.pairs'.
assignments <- cyclone(sce_obj, mm.pairs, gene.names=rownames(sce_obj))
saveRDS(assignments,"../data_output/osca_objects/cellcycle_assignments_rgc_all_litters.rds")
```


```{r}
plot(assignments$score$G1, assignments$score$G2M,
    xlab="G1 score", ylab="G2/M score", pch=16)

```

```{r}
assignments <- readRDS("../data_output/osca_objects/cellcycle_assignments_rgc_all_litters.rds")

seurat_obj$G1 <- assignments$normalized.scores$G1
seurat_obj$S <- assignments$normalized.scores$S
seurat_obj$G2M <- assignments$normalized.scores$G2M
seurat_obj$phases <- assignments$phases

```

```{r}
FeaturePlot(seurat_obj,features = "G1")
FeaturePlot(seurat_obj,features = "G2M")
FeaturePlot(seurat_obj,features = "S")

```

```{r}

pdf("../data_output/20230907_rgc_both_litters_umap_cell_cycle.pdf")
FeaturePlot(seurat_obj,features = "G1")
FeaturePlot(seurat_obj,features = "G2M")
FeaturePlot(seurat_obj,features = "S")
DimPlot(seurat_obj,group.by  = "phases") + theme(aspect.ratio = 1)
DimPlot(seurat_obj,group.by  = "phases", split.by = "geno") + theme(aspect.ratio = 1)
dev.off()
```

```{r}
table(seurat_obj$seurat_clusters,paste(seurat_obj$phases,seurat_obj$geno,sep = "_"))
```


# Differential expression
```{r}
library(scran)
library(scater)
sce_obj <- SingleCellExperiment(assays = list(counts = seurat_obj@assays$RNA@counts), 
                           colData = DataFrame(seurat_obj@meta.data))
summed <- aggregateAcrossCells(sce_obj, 
    id=colData(sce_obj)[,c("seurat_clusters","sample")])

summed$batch <- substr(summed$sample,3,3)
# Creating up a DGEList object for use in edgeR:
library(edgeR)

de_results <- pseudoBulkDGE(summed,
                            sample = summed$sample,
    label=summed$seurat_clusters,
    design= ~ batch + geno,
    contrast = "genoXP",
    condition=summed$geno
)
```

```{r}
ambient_matrix <- readRDS(file = "../data_output/osca_objects/ambient_matrix.Rds")

```


```{r}

  summed <- summed[rownames(summed) %in% rownames(ambient_matrix),]
  
  ambient_matrix <- ambient_matrix[rownames(summed), ]


maxambients_per_clust <- map(unique(summed$seurat_clusters),function(.x, summed, ambient){
  current_summed <- summed[,summed$seurat_clusters == .x]
  
  current_max_ambient <- DropletUtils::ambientContribMaximum(counts(current_summed), 
                                               ambient_matrix, mode="proportion")
  
  
  
  currentcontamination <- rowMeans(current_max_ambient, na.rm=TRUE)
  return(currentcontamination)
}, summed = summed, ambient = ambient_matrix)


names(maxambients_per_clust) <- unique(summed$seurat_clusters)
```

```{r}
is_de <- decideTestsPerLabel(de_results, threshold=0.05)
summarizeTestsPerLabel(is_de)

```


```{r}
lognormsummed <- summed
sizeFactors(lognormsummed) <- NULL
lognormsummed <- logNormCounts(lognormsummed)

```

```{r}
library("org.Mm.eg.db")

de_results_df_list <- lapply(names(de_results),
      function(clust,de_df, ambient_mat) {
        x <- de_df[[clust]]
        contamination <- ambient_mat[[clust]]
        x <- as.data.frame(x)
        x$cluster <- clust
        x$contamination <- contamination[rownames(x)]
        x <-rownames_to_column(x,var = "ENSEMBL")
        
        return(x)
      }, de_df = de_results, ambient_mat = maxambients_per_clust)


de_results_df <- do.call(rbind,de_results_df_list)

ens2symbol <- clusterProfiler::bitr(de_results_df$ENSEMBL,fromType = "ENSEMBL",toType = "SYMBOL",OrgDb = org.Mm.eg.db)

de_results_df <- left_join(de_results_df, ens2symbol,by= "ENSEMBL")
de_results_df <- drop_na(de_results_df)
```

```{r}
de_results_df_filt <- de_results_df[de_results_df$FDR < 0.05 & de_results_df$contamination < 0.05,]
```
```{r}
write.csv(de_results_df,file = "../data_output/surat_objects/rgc_clusters_deg_all.csv")
```

```{r}
write.csv(de_results_df_filt,file = "../data_output/surat_objects/fzd8_clusters_deg_fdr05_contamination_05.csv")
```

```{r}
pdf("../data_output/20230907_rgc_both_litters_diffexpression.pdf")

de_results_df %>%
  ggplot(aes(logCPM, logFC, color = FDR < 0.05 & contamination < 0.05)) +
  geom_point(size = 0.5, pch = 1) +
  scale_color_manual(values = c("TRUE" = "red","FALSE"="darkgray")) +
  theme_light() +
  theme(aspect.ratio = 1) +
  facet_wrap(~cluster) +
  ggrepel::geom_text_repel(mapping = aes(label = SYMBOL),
                           data = de_results_df %>%
                             filter(FDR < 0.05 & contamination < 0.05),
                           color = "black",
                           max.overlaps = 30,
                           size = 3) 
dev.off()
```
Show only WNT signaling genes
```{r}

mmu_kegg <- clusterProfiler::download_KEGG(species ="mmu")

wnt_pathway <- mmu_kegg$KEGGPATHID2NAME[str_detect(mmu_kegg$KEGGPATHID2NAME$to,"W"), ]

wnt_pathway_entrez_genes <- mmu_kegg$KEGGPATHID2EXTID %>%
  filter(from == wnt_pathway$from )

wnt_pathway_genes <- clusterProfiler::bitr(wnt_pathway_entrez_genes$to,OrgDb = org.Mm.eg.db,fromType = "ENTREZID",toType=c("ENSEMBL","SYMBOL"))


```

```{r}
de_results_df %>%
  ggplot(aes(logCPM, logFC, color = FDR < 0.05 &contamination < 0.05 & ENSEMBL %in% wnt_pathway_genes$ENSEMBL )) +
  geom_point(size = 0.5, pch = 1) +
  scale_color_manual(values = c("TRUE" = "red","FALSE"="darkgray")) +
  theme_light() +
  theme(aspect.ratio = 1,legend.position = "none") +
  facet_wrap(~cluster) +
  ggrepel::geom_text_repel(mapping = aes(label = SYMBOL),
                           data = de_results_df %>%
                             filter(FDR < 0.05 & contamination < 0.05 & ENSEMBL %in% wnt_pathway_genes$ENSEMBL),
                           color = "black",
                           max.overlaps = 30,
                           size = 3) 

```


```{r}
DefaultAssay(seurat_obj) <- "RNA"

seurat_obj <- NormalizeData(seurat_obj)

Uba52 <- bitr("Zic4",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

FeaturePlot(object = seurat_obj,features = Uba52,split.by = "geno") +
  theme(aspect.ratio = 1)

VlnPlot(object = seurat_obj, features = Uba52,split.by = "geno") +
  theme(aspect.ratio = 0.75)
```

```{r}

H3c2 <- bitr("Lix1",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb = org.Mm.eg.db)$ENSEMBL

FeaturePlot(object = seurat_obj,features = H3c2,split.by = "geno") +
  theme(aspect.ratio = 1)

VlnPlot(object = seurat_obj, features = H3c2,split.by = "geno") +
  theme(aspect.ratio = 0.75)
```
