---
title: "cell_cycle_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(scater)
library(scran)
library(edgeR)
library(Seurat)
library(tidyverse)
options(Seurat.memsafe = TRUE)

```

## Cell cycle analysis 

Here I follow the chapter from: http://bioconductor.org/books/3.14/OSCA.advanced/cell-cycle-assignment.html

## Aim

Cell cycle analysis is assigning a cell cycle phase (G1, S, G2, etc) to each cell. This information has been used as a nuissance factor with the assumption that differences in cell cycle stage between cells of the same cell type are not interesting, and can drive unwanted cluster separation. 

In addition, in this case we are most interested in whether there are changes in cell cycle stages between experimental conditions.

## Approach

Using the cyclone classifier:
- Uses a reference dataset, precomputed for mouse. 
- This dataset has cells in different cell cycle stages, finds DE genes between cell cycle stages, chooses these as markers. 
- Then, the sign of the markers in the test dataset are used to assign cells to stages.

## Prepare data 
```{r}
sce_obj <- readRDS("../data_output/surat_objects/run_8nov21/Mm_merged_mito_mito_integrated.rds")

DefaultAssay(sce_obj) <- "RNA"


sce_obj <- SingleCellExperiment(assays = list(counts = sce_obj@assays$RNA@counts), 
                           colData = DataFrame(sce_obj@meta.data))


```

## Run cyclone
```{r eval = FALSE}
set.seed(100)
library(scran)
mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", 
    package="scran"))

# Using Ensembl IDs to match up with the annotation in 'mm.pairs'.
assignments <- cyclone(sce_obj, mm.pairs, gene.names=rownames(sce_obj))
plot(assignments$score$G1, assignments$score$G2M,
    xlab="G1 score", ylab="G2/M score", pch=16)
saveRDS(assignments,"../data_output/osca_objects/cellcycle_assignments.rds")
```
## Explore cell cycle scores in clusters
```{r}
assignments <- readRDS("../data_output/osca_objects/cellcycle_assignments.rds")
sce_obj <- readRDS("../data_output/surat_objects/run_8nov21/Mm_merged_mito_mito_integrated.rds")
```
```{r}
sce_obj$G1 <- assignments$normalized.scores$G1
sce_obj$S <- assignments$normalized.scores$S
sce_obj$G2M <- assignments$normalized.scores$G2M
sce_obj$phases <- assignments$phases
```

```{r}
FeaturePlot(sce_obj,features = "G1")
```

```{r}
FeaturePlot(sce_obj,features = "G2M")
```

```{r}
FeaturePlot(sce_obj,features = "S")

```

```{r}
DimPlot(sce_obj,group.by  = "phases")

```

```{r}
cells_by_cycle <- table(sce_obj$phases,sce_obj$sample)
```

```{r}
(cells_by_cycle[2,]/cells_by_cycle[1,]) %>%
  enframe() %>%
  ggplot(aes(name,value)) +
  geom_col()+
  theme(aspect.ratio = 1) +
  ylab("G2-M / G1") +
  xlab(NULL)

(cells_by_cycle[3,]/cells_by_cycle[1,]) %>%
  enframe() %>%
  ggplot(aes(name,value)) +
  geom_col() +
  theme(aspect.ratio = 1)

```
By cluster
```{r}
counts_per_cluster <- table(sce_obj$phases,paste(sce_obj$seurat_clusters, sce_obj$sample))
g2_g1_percluster<- counts_per_cluster["G2M",] / counts_per_cluster["G1",]

```

```{r}
g2_g1_percluster_df <- g2_g1_percluster %>%
  enframe(name = "sample_cluster",value = "g2m/g1") %>%
  mutate(cluster = str_split(sample_cluster, " ") %>% map_chr(1),
         sample = str_split(sample_cluster, " ") %>% map_chr(2))
  
  
```
```{r}
g2_g1_percluster_df %>%
  ggplot(aes(sample, `g2m/g1`)) +
  geom_col() +
  facet_wrap(vars(cluster))
```

# Check cyclin genes
```{r}
library(clusterProfiler)
library(org.Mm.eg.db)
gene_names <- bitr(rownames(sce_obj),fromType = "ENSEMBL",toType = "SYMBOL", OrgDb = org.Mm.eg.db)

cyclin_genes <- gene_names[grepl("^Ccn[abde][0-9]$", gene_names$SYMBOL),]

cyclin_genes
```
G2 markers

Ccnb1
```{r}
FeaturePlot(sce_obj,features = "ENSMUSG00000041431")

```
Ccna2 (Also S)
```{r}
FeaturePlot(sce_obj,features = "ENSMUSG00000027715")

```

G1 markers

Cyclin D
```{r}
FeaturePlot(sce_obj,features = "ENSMUSG00000000184")

```
Cyclin E

G1/S
```{r}
FeaturePlot(sce_obj,features = "ENSMUSG00000028212")

```


