---
title: "cell_cycle_tutorial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cell cycle analysis 

Here I follow the chapter from: http://bioconductor.org/books/3.14/OSCA.advanced/cell-cycle-assignment.html

## Aim

Cell cycle analysis is assigning a cell cycle phase (G1, S, G2, etc) to each cell. This information has been used as a nuissance factor with the assumption that differences in cell cycle stage between cells of the same cell type are not interesting, and can drive unwanted cluster separation. 

In addition, in this case we are most interested in whether there are changes in cell cycle stages between experimental conditions.

## Approaches
- Cyclin genes
- Reference profiles
- Classifier

## Prepare example dataset

The example dataset:

"the 416B dataset, which is known to contain actively cycling cells after oncogene induction"

This is the script to prepare the example dataset.
```{r}
#--- loading ---#
library(scRNAseq)
sce.416b <- LunSpikeInData(which="416b") 
sce.416b$block <- factor(sce.416b$block)

#--- gene-annotation ---#
library(AnnotationHub)
ens.mm.v97 <- AnnotationHub()[["AH73905"]]
rowData(sce.416b)$ENSEMBL <- rownames(sce.416b)
rowData(sce.416b)$SYMBOL <- mapIds(ens.mm.v97, keys=rownames(sce.416b),
    keytype="GENEID", column="SYMBOL")
rowData(sce.416b)$SEQNAME <- mapIds(ens.mm.v97, keys=rownames(sce.416b),
    keytype="GENEID", column="SEQNAME")

library(scater)
rownames(sce.416b) <- uniquifyFeatureNames(rowData(sce.416b)$ENSEMBL, 
    rowData(sce.416b)$SYMBOL)

#--- quality-control ---#
mito <- which(rowData(sce.416b)$SEQNAME=="MT")
stats <- perCellQCMetrics(sce.416b, subsets=list(Mt=mito))
qc <- quickPerCellQC(stats, percent_subsets=c("subsets_Mt_percent",
    "altexps_ERCC_percent"), batch=sce.416b$block)
sce.416b <- sce.416b[,!qc$discard]

#--- normalization ---#
library(scran)
sce.416b <- computeSumFactors(sce.416b)
sce.416b <- logNormCounts(sce.416b)

#--- variance-modelling ---#
dec.416b <- modelGeneVarWithSpikes(sce.416b, "ERCC", block=sce.416b$block)
chosen.hvgs <- getTopHVGs(dec.416b, prop=0.1)

#--- batch-correction ---#
library(limma)
assay(sce.416b, "corrected") <- removeBatchEffect(logcounts(sce.416b), 
    design=model.matrix(~sce.416b$phenotype), batch=sce.416b$block)

#--- dimensionality-reduction ---#
sce.416b <- runPCA(sce.416b, ncomponents=10, subset_row=chosen.hvgs,
    exprs_values="corrected", BSPARAM=BiocSingular::ExactParam())

set.seed(1010)
sce.416b <- runTSNE(sce.416b, dimred="PCA", perplexity=10)

#--- clustering ---#
my.dist <- dist(reducedDim(sce.416b, "PCA"))
my.tree <- hclust(my.dist, method="ward.D2")

library(dynamicTreeCut)
my.clusters <- unname(cutreeDynamic(my.tree, distM=as.matrix(my.dist),
    minClusterSize=10, verbose=0))
colLabels(sce.416b) <- factor(my.clusters)

```

## Using the cyclins

G1 - Cyclin D peaks
G1/S - Cyclin E
S - Cyclin A
S/G2 - Cyclin A
G2/Mitosis - Cyclin B

```{r}
library(scater)
cyclin.genes <- grep("^Ccn[abde][0-9]$", rowData(sce.416b)$SYMBOL)
cyclin.genes <- rownames(sce.416b)[cyclin.genes]
cyclin.genes

```

```{r}
plotHeatmap(sce.416b, order_columns_by="label", 
    cluster_rows=FALSE, features=sort(cyclin.genes))

```

## Reference profile

I am not going to try this because I don't know if there is something like this for mouse cells. 

## Cyclone classifier

How it works:

Also uses a reference dataset, but precomputed for mouse are available. 
This dataset has cells in different cell cycle stages, finds DE genes between cell cycle stages, chooses these as markers. 

Then, the sign of the markers in the test dataset are used to assign cells to stages.

```{r}
set.seed(100)
library(scran)
mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", 
    package="scran"))

# Using Ensembl IDs to match up with the annotation in 'mm.pairs'.
assignments <- cyclone(sce.416b, mm.pairs, gene.names=rowData(sce.416b)$ENSEMBL)
plot(assignments$score$G1, assignments$score$G2M,
    xlab="G1 score", ylab="G2/M score", pch=16)

```
```{r}
plotReducedDim(sce.416b,dimred = "TSNE", colour_by =  DataFrame(assignments$normalized.scores$G1))
plotReducedDim(sce.416b,dimred = "TSNE", colour_by =  DataFrame(assignments$normalized.scores$G2))

plotReducedDim(sce.416b,dimred = "TSNE", colour_by =  DataFrame(assignments$normalized.scores$S))

plotReducedDim(sce.416b,dimred = "TSNE", colour_by =  DataFrame(assignments$phases))

```


