---
title: "diff_abundance.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The aim of differential abundance is to see if there are changes in cell type composition.
I will be following this chapter from OSCA: http://bioconductor.org/books/3.14/OSCA.multisample/differential-abundance.html


Read in gastrulation dataset:
```{r}
merged <- readRDS("../data_output/mouse_gastru_merged.rds")

```

Compute abundances:
```{r}
abundances <- table(merged$celltype.mapped, merged$sample) 
abundances <- unclass(abundances) 
head(abundances)

```

```{r}
library(edgeR)

extra.info <- colData(merged)[match(colnames(abundances), merged$sample),]
y.ab <- DGEList(abundances, samples=extra.info)
y.ab

```

Filter low abundance cell types.

```{r}
keep <- filterByExpr(y.ab, group=y.ab$samples$tomato)
y.ab <- y.ab[keep,]
summary(keep)

```

Unlike differential expression, they don't recommend normalizing by composition.

Design matrix
```{r}
design <- model.matrix(~factor(pool) + factor(tomato), y.ab$samples)

```

Estimate dispersion without trend (no enough points for reliable estimation of trended dispersion).
```{r}
y.ab <- estimateDisp(y.ab, design, trend="none")

plotBCV(y.ab, cex=1)

```
QL dispersion
```{r}
fit.ab <- glmQLFit(y.ab, design, robust=TRUE, abundance.trend=FALSE)

```

Test differences
```{r}
res <- glmQLFTest(fit.ab, coef=ncol(design))
summary(decideTests(res))

topTags(res)

```




