---
title: "diff_abundance_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)

library(Seurat)
library(tidyverse)
library(edgeR)
library(scran)
```


## Aim

The aim of differential abundance is to test for changes in cell type composition between WT and XP conditions: http://bioconductor.org/books/3.14/OSCA.multisample/differential-abundance.html


## Analysis overview

1. Count number of cells per cluster per condition per replicate 
2. Test for differential abundance using edgeR

## Pre-processing and summarizing counts
```{r}
sce_obj <- as.SingleCellExperiment(readRDS("../data_output/surat_objects/run_8nov21/Mm_merged_mito_mito_integrated.rds"))


abundances <- table(sce_obj$seurat_clusters, sce_obj$sample) 
abundances <- unclass(abundances) 


extra_info <- colData(sce_obj)[match(colnames(abundances), sce_obj$sample),]
extra_info$batch <- factor(substr(extra_info$sample,3,3))
extra_info$geno <- factor(extra_info$geno)
y_ab <- DGEList(abundances, samples=extra_info)

```


## Testing for DA

Design matrix
```{r}

design <- model.matrix(~ batch + geno, y_ab$samples)

```

Estimate dispersions
```{r}
y_ab <- calcNormFactors(y_ab)
y_ab <- estimateDisp(y_ab, design, trend="none")

plotBCV(y_ab, cex=1)

```

```{r}
fit_ab <- glmQLFit(y_ab, design, robust=TRUE, abundance.trend=FALSE)

```

```{r}
res <- glmQLFTest(fit_ab, coef=ncol(design))
summary(decideTests(res))

```


There seem not to be any differentially abundant cell types (or these are just too variable between batches that it's not really reliable).
```{r}
topTags(res,sort.by = "PValue", n = 5)$table
```


## Top 5 by pvalue: 
```{r}
colData(sce_obj) %>%
  as_data_frame() %>%
  mutate(seurat_clusters = ordered(seurat_clusters,levels =as.character(sort(as.numeric(unique(seurat_clusters)))) ),
         is_da = seurat_clusters %in% rownames(topTags(res,sort.by = "PValue", n = 5))) %>%
  filter(is_da) %>%
  ggplot(aes(seurat_clusters, fill =  sample,)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(type = "div")

```

## Use DA analysis to identify batch specific clusters

Design matrix
```{r}

design <- model.matrix(~ batch, y_ab$samples)
y_ab <- calcNormFactors(y_ab)
y_ab <- estimateDisp(y_ab, design, robust = TRUE, trend="none")

```

```{r}
plotBCV(y_ab, cex=1)
```


```{r}
fit_ab <- glmQLFit(y_ab, design, robust=TRUE, abundance.trend=FALSE)

res <- glmQLFTest(fit_ab, coef=ncol(design))

```

```{r}
summary(decideTests(res,p.value = 0.001,lfc = 0.5))
```


```{r}
topTags(res,sort.by = "PValue", n = Inf)$table %>%
  ggplot(aes(logFC, - 10 * log10(FDR))) +
  geom_point() +
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = - 10 * log10(0.001)) +
  geom_vline(xintercept = -3)
```
Clusters to remove:

```{r}
res$table[decideTests(res,lfc = 3,p.value = 1)<0,]
```

```{r eval = FALSE, include = FALSE}
saveRDS(res$table,"../data_output/osca_objects/batch_biased_clusters.rds")
```


```{r}
clust_to_remove <- rownames(res$table[decideTests(res,lfc = 3,p.value = 1)<0,])
```


```{r}
colData(sce_obj) %>%
  as_data_frame() %>%
  mutate(seurat_clusters = ordered(seurat_clusters,levels =as.character(sort(as.numeric(unique(seurat_clusters)))) ),
         is_da = seurat_clusters %in% clust_to_remove) %>%
  filter(is_da) %>%
  ggplot(aes(seurat_clusters, fill =  sample,)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(type = "div")

```


## Re-do analysis without problematic clusters

## Testing for DA

```{r}
y_ab <- y_ab[! rownames(y_ab) %in% clust_to_remove,]
```


Design matrix
```{r}

design <- model.matrix(~ batch + geno, y_ab$samples)

```

Estimate dispersions
```{r}
y_ab <- calcNormFactors(y_ab)
y_ab <- estimateDisp(y_ab, design,robust = TRUE, trend="none")

plotBCV(y_ab, cex=1)

```

```{r}
fit_ab <- glmQLFit(y_ab, design, robust=TRUE, abundance.trend=FALSE)

```

```{r}
res <- glmQLFTest(fit_ab, coef=ncol(design))
summary(decideTests(res))

```


There seem not to be any differentially abundant cell types (or these are just too variable between batches that it's not really reliable).
```{r}
topTags(res,sort.by = "PValue", n = 5)$table
```


## Top 5 by pvalue: 
```{r}
colData(sce_obj) %>%
  as_data_frame() %>%
  mutate(seurat_clusters = ordered(seurat_clusters,levels =as.character(sort(as.numeric(unique(seurat_clusters)))) ),
         is_da = seurat_clusters %in% rownames(topTags(res,sort.by = "PValue", n = 5))) %>%
  filter(is_da) %>%
  ggplot(aes(seurat_clusters, fill =  sample,)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(type = "div")

```
