---
title: "milo_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(miloR)
library(SingleCellExperiment)
library(scater)
library(dplyr)
library(patchwork)
library(clusterProfiler)
library(org.Mm.eg.db)
library(Seurat)

```

# Test for differential abundance

# Aim

Milo detects changes in composition from different experiments, with replication and complex designs (like batch and experimental conditions).


## Step 1: Build Milo object from seurat object with four samples

Read seurat object. 
```{r}
seurat_obj <- readRDS("../data_output/surat_objects/run_8nov21/Mm_merged_mito_mito_integrated.rds")
```

Filter malat1 cells
```{r}
malat_ensembl <- bitr("Malat1",fromType = "SYMBOL",toType = "ENSEMBL",OrgDb =  org.Mm.eg.db)$ENSEMBL
DefaultAssay(seurat_obj) <- "RNA"

seurat_obj$percent.malat1 <- PercentageFeatureSet(seurat_obj,features = malat_ensembl)

seurat_obj <- seurat_obj[,seurat_obj$percent.malat1 < 5]


```

```{r}
DefaultAssay(seurat_obj) <- "integrated"

seurat_obj <- RunPCA(seurat_obj,npcs = 100)
seurat_obj <- RunUMAP(seurat_obj,dims = 1:50)

```


Build Milo object

```{r}
sce_obj <- as.SingleCellExperiment(seurat_obj)

milo_obj <- Milo(sce_obj)
```

## Step 2: Build kNN graph from integrated PCA space

```{r}
milo_obj <- buildGraph(milo_obj, k = 30, d = 50, reduced.dim = "PCA")

```

## Step 3: Get representative neighbourhoods

```{r}
milo_obj <- makeNhoods(milo_obj, prop = 0.1, k = 30, d = 50, refined = TRUE, reduced_dims = "PCA")

```

```{r}
plotNhoodSizeHist(milo_obj)
```
## Step 4: Count number of cells per neighbourhood per condition

```{r}
milo_obj <- countCells(milo_obj, meta.data = as.data.frame(colData(milo_obj)), sample="sample")

```

## Step 5: Build design matrix 

```{r}
design <- data.frame(colData(milo_obj))[,c("sample", "geno")]

design <- design %>%
  mutate(batch = as.factor(substr(sample,3,3))) 

design <- distinct(design)

rownames(design) <- design$sample

```

## Step 6: Computing neighbourhood connectivity
```{r}
milo_obj <- calcNhoodDistance(milo_obj, d=50, reduced.dim = "PCA")


```

```{r}
da_results <- testNhoods(milo_obj, design = ~ batch + geno, design.df = design, reduced.dim="PCA")

```

```{r}
da_results %>%
  arrange(SpatialFDR) %>%
  head() 

```
```{r}
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)

```
```{r}
ggplot(da_results, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)

```
```{r}
milo_obj <- buildNhoodGraph(milo_obj)

## Plot single-cell UMAP
umap_pl <- plotReducedDim(milo_obj, dimred = "UMAP", colour_by="geno") +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo_obj, da_results, layout="UMAP",alpha=0.1) 
  
umap_pl + nh_graph_pl +
  plot_layout(guides="collect")

```

