---
title: "milo_tutorial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Aim

Milo detects changes in composition from different experiments, with replication.

# Overview

The point is to not aggregate cells into discrete clusters, but rather dectect changes in composition more generally.

There are some problems with aggregating cells into discrete clusters and then calculating DA.

- It builds a kNN graph.
-

# Libraries
```{r}
library(miloR)
library(SingleCellExperiment)
library(scater)
library(dplyr)
library(patchwork)

```


# Load test data
```{r}
data("sim_trajectory", package = "miloR")

## Extract SingleCellExperiment object
traj_sce <- sim_trajectory[['SCE']]

## Extract sample metadata to use for testing
traj_meta <- sim_trajectory[["meta"]]

## Add metadata to colData slot
colData(traj_sce) <- DataFrame(traj_meta)

```

# Pre-processing


Build kNN graph from PCA distances.

First, do PCA. 

And UMAP for visualization.

```{r}
logcounts(traj_sce) <- log(counts(traj_sce) + 1)
traj_sce <- runPCA(traj_sce, ncomponents=30)
traj_sce <- runUMAP(traj_sce)

plotUMAP(traj_sce)

```

Build 
```{r}
traj_milo <- Milo(traj_sce)
reducedDim(traj_milo, "UMAP") <- reducedDim(traj_sce, "UMAP")

```

```{r}
traj_milo <- buildGraph(traj_milo, k = 10, d = 30)

```

Make neighbourhoods

```{r}
traj_milo <- makeNhoods(traj_milo, prop = 0.1, k = 10, d=30, refined = TRUE)

```

How big are the neighbourhoods?
```{r}
plotNhoodSizeHist(traj_milo)

```

Count cells per neighbourhood

```{r}
traj_milo <- countCells(traj_milo, meta.data = data.frame(colData(traj_milo)), sample="Sample")
```

```{r}
head(nhoodCounts(traj_milo))

traj_design <- data.frame(colData(traj_milo))[,c("Sample", "Condition")]
traj_design <- distinct(traj_design)


rownames(traj_design) <- colnames(traj_milo@nhoodCounts)
```


```{r}
traj_milo <- calcNhoodDistance(traj_milo, d=30)

```

```{r}
da_results <- testNhoods(traj_milo, design = ~ Condition, design.df = traj_design)

traj_milo
```

```{r}
da_results %>%
  arrange(- SpatialFDR) %>%
  head()

```
```{r}
traj_milo <- buildNhoodGraph(traj_milo)

```
```{r}
plotUMAP(traj_milo) + plotNhoodGraphDA(traj_milo, da_results, alpha=0.05) +
  plot_layout(guides="collect")

```

# Mouse gastrulation

```{r}
library(miloR)
library(SingleCellExperiment)
library(scater)
library(scran)
library(dplyr)
library(patchwork)
library(MouseGastrulationData)

```

```{r}
select_samples <- c(2,  3,  6, 4, #15,
                    # 19, 
                    10, 14#, 20 #30
                    #31, 32
                    )
embryo_data = EmbryoAtlasData(samples = select_samples)
embryo_data

```


```{r}
embryo_data <- embryo_data[,apply(reducedDim(embryo_data, "pca.corrected"), 1, function(x) !all(is.na(x)))]
embryo_data <- runUMAP(embryo_data, dimred = "pca.corrected", name = 'umap')

plotReducedDim(embryo_data, colour_by="stage", dimred = "umap") 

```

```{r}
embryo_milo <- Milo(embryo_data)
embryo_milo <- buildGraph(embryo_milo, k = 30, d = 30, reduced.dim = "pca.corrected")
embryo_milo <- makeNhoods(embryo_milo, prop = 0.1, k = 30, d=30, refined = TRUE, reduced_dims = "pca.corrected")
plotNhoodSizeHist(embryo_milo)
embryo_milo <- countCells(embryo_milo, meta.data = as.data.frame(colData(embryo_milo)), sample="sample")

embryo_design <- data.frame(colData(embryo_milo))[,c("sample", "stage", "sequencing.batch")]

## Convert batch info from integer to factor
embryo_design$sequencing.batch <- as.factor(embryo_design$sequencing.batch) 
embryo_design <- distinct(embryo_design)
rownames(embryo_design) <- embryo_design$sample

embryo_milo <- calcNhoodDistance(embryo_milo, d=30, reduced.dim = "pca.corrected")

da_results <- testNhoods(embryo_milo, design = ~ sequencing.batch + stage, design.df = embryo_design, reduced.dim="pca.corrected")

head(da_results)

```

```{r}
da_results %>%
  arrange(SpatialFDR) %>%
  head() 

```
Inspecting results

```{r}
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)

```


