---
title: "scMm_analysis_aj_with_mito"
output: html_document
---

Follow this code as closely as possible to reproduce the results.

Get these files:

Mm_merged.rds

Mm_umap_integrated.rds
Mm_umap_integrated_markers.rds


Mm_umap_integrated_subset.rds
Mm_umap_integrated_markers_subset.rds


```{r}
library(Matrix)
library(dplyr)
library(Seurat)
library(tinytex)
library(patchwork)
library(ggplot2)
options(Seurat.memsafe = TRUE)
getwd()
```


```{r}
matrix_dir = "../data_raw/scmm_project/G_A1/outs/filtered_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
WT1 <- readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(WT1) = barcode.names$V1
rownames(WT1) = feature.names$V1
```

```{r}
matrix_dir = "../data_raw/scmm_project/G_B1/outs/filtered_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
XP1 <- readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(XP1) = barcode.names$V1
rownames(XP1) = feature.names$V1
```

```{r}
matrix_dir = "../data_raw/scmm_project/G_C2/outs/filtered_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
WT2 <- readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(WT2) = barcode.names$V1
rownames(WT2) = feature.names$V1
```

```{r}
matrix_dir = "../data_raw/scmm_project/G_D2/outs/filtered_feature_bc_matrix/"
barcode.path <- paste0(matrix_dir, "barcodes.tsv.gz")
features.path <- paste0(matrix_dir, "features.tsv.gz")
matrix.path <- paste0(matrix_dir, "matrix.mtx.gz")
XP2 <- readMM(file = matrix.path)
feature.names = read.delim(features.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path,
                           header = FALSE,
                           stringsAsFactors = FALSE)
colnames(XP2) = barcode.names$V1
rownames(XP2) = feature.names$V1
```

```{r convert files to seurat}
WT1_seurat <- CreateSeuratObject(counts = WT1, project = "WT1", min.cells = 3, min.features = 200)
WT2_seurat <- CreateSeuratObject(counts = WT2, project = "WT2", min.cells = 3, min.features = 200)
XP1_seurat <- CreateSeuratObject(counts = XP1, project = "XP1", min.cells = 3, min.features = 200)
XP2_seurat <- CreateSeuratObject(counts = XP2, project = "XP2", min.cells = 3, min.features = 200)

WT1_seurat$sample <- "WT1"
WT2_seurat$sample <- "WT2"
XP1_seurat$sample <- "XP1"
XP2_seurat$sample <- "XP2"

WT1_seurat$geno <- "WT"
WT2_seurat$geno <- "WT"
XP1_seurat$geno <- "XP"
XP2_seurat$geno <- "XP"
```

##merging the seurat objects if interested 
```{r merged seurat object}
Mm.merged <- merge(WT1_seurat, y = c(WT2_seurat, XP1_seurat, XP2_seurat), add.cell.ids = c("WT1", "WT2", "XP1", "XP2"), project = "Mm_merged_rep")
Mm.merged.subset <- merge(WT2_seurat, y = c(XP2_seurat), add.cell.ids = c("WT2", "XP2"), project = "Mm_merged_rep")
```

```{r}
Mm.merged$group <- factor(Mm.merged$sample, levels = c("WT1", "WT2", "XP1", "XP2"))
Mm.merged.subset$group <- factor(Mm.merged$sample, levels = c("WT2", "XP2"))
```

```{r}
saveRDS(Mm.merged, file = "../data_output/surat_objects/Mm_merged_rep.rds")
saveRDS(Mm.merged.subset, file = "../data_output/surat_objects/Mm_merged_subset_rep.rds")
```

###PART 2 

#load from drive
```{r}
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged.rds")
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged_subset.rds")
```
##put in mito at this point
```{r}
Mm.merged[["percent.mt"]] <- PercentageFeatureSet(Mm.merged, pattern = "^MT-")
Mm.merged.subset[["percent.mt"]] <- PercentageFeatureSet(Mm.merged.subset, pattern = "^MT-")

```
##new merge copy and reordering the columns
```{r}
Mm.merged.ordered <- Mm.merged
my_levels <- c("WT1","XP1","WT2","XP2")
Idents(Mm.merged.ordered) <- factor(Idents(Mm.merged.ordered), levels= my_levels)


VlnPlot(Mm.merged.ordered, features = c("nFeature_RNA", "nCount_RNA"), pt.size = 0, ncol = 2)
```

```{r}
VlnPlot(Mm.merged.ordered, features = c("nFeature_RNA"), pt.size = 1E-100)
VlnPlot(Mm.merged.ordered, features = c("nCount_RNA"), pt.size = 1E-100)
```

```{r}
plot1 <- FeatureScatter(Mm.merged.ordered, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Mm.merged.ordered, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

##Subsetting the data for all samples
```{r}
Mm.merged <- subset(Mm.merged.ordered, subset = nFeature_RNA > 500 & nFeature_RNA < 7500 & nCount_RNA < 50000 & percent.mt < 5)
VlnPlot(Mm.merged, features = c("nFeature_RNA", "nCount_RNA"), pt.size = 0, ncol = 2)
```

```{r}
dim(Mm.merged)
```

```{r}
saveRDS(Mm.merged, file = "../data_output/surat_objects/Mm_merged_mito_rep.rds")
saveRDS(Mm.merged.subset, file = "../data_output/surat_objects/Mm_merged_mito_subset_rep.rds")
```
```{r}
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged_mito.rds")
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged_mito_subset.rds")
```

```{r}
library(dplyr)
library(magrittr)
library(tidyr)
library(tidyverse)
library(Seurat)
library(sctransform)
```

```{r}
Mm.merged.transformed <- SCTransform(Mm.merged, variable.features.n = 5000, vars.to.regress = "percent.mt", verbose = FALSE)
Mm.merged.subset.transformed <- SCTransform(Mm.merged.subset, variable.features.n = 5000, vars.to.regress = "percent.mt", verbose = FALSE)

```
```{r}
saveRDS(Mm.merged.transformed, file = "../data_output/surat_objects/Mm_merged_sct_rep.rds")
saveRDS(Mm.merged.subset.transformed, file = "../data_output/surat_objects/Mm_merged_sct_subset_rep.rds")
```

```{r}
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged_sct.rds")
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged_sct_subset.rds")

Mm.merged <- Mm.merged.transformed
Mm.merged.subset <- Mm.merged.subset.transformed
```

# These are now standard steps in the Seurat workflow for visualization and clustering

##PCA analysis with 200 pcs to test which is best but will try 300
```{r}
Mm.merged <- RunPCA(object = Mm.merged, npcs = 300, features = VariableFeatures(object = Mm.merged))
Mm.merged.subset <- RunPCA(object = Mm.merged.subset, npcs = 300, features = VariableFeatures(object = Mm.merged.subset))

```



```{r}
saveRDS(Mm.merged, file = "../data_output/surat_objects/Mm_merged_sct_pca_rep.rds")

saveRDS(Mm.merged.subset, file = "../data_output/surat_objects/Mm_merged_sct_pca_subset_rep.rds")
```
```{r}
Mm.merged <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_merged_sct_pca.rds")
Mm.merged <- readRDS("Mm_merged_sct_pca_subset_rep.rds")
```

```{r}
library("dplyr")
library("Seurat")
```
##Print genes
```{r}
print(Mm.merged[["pca"]], dims = 1:5, nfeatures = 5)
```
## view important PCAs
```{r}
VizDimLoadings(Mm.merged, dims = 105, reduction = "pca")
```
##view the PCA
```{r}
DimPlot(Mm.merged, reduction = "pca")
```
##Heatmap and in multiple dimensions
```{r}
DimHeatmap(Mm.merged, dims = 195:200, cells = 2000, balanced = TRUE)
```

```{r}
ElbowPlot(Mm.merged, ndims = 300)
```

##125 is good 24,1.0 has 36 (good option)
1.5 is 35
100 PCs good, 1.1 25 clusters
```{r}
Mm_neighbor <- FindNeighbors(object = Mm.merged, dims = 1:100)
Mm_clusters <- FindClusters(object = Mm_neighbor, resolution = 1.1)
Mm_umap <- RunUMAP(Mm_clusters, dims = 1:100)
DimPlot(Mm_umap, reduction = "umap", label = TRUE)
DimPlot(Mm_umap, reduction = "umap", group.by = "group", label = TRUE, repel = TRUE)
DimPlot(Mm_umap, reduction = "umap", group.by = "geno", label = TRUE, repel = TRUE)
dim(Mm_umap)
```

Same but for subset 
```{r}
Mm.merged <- FindNeighbors(object = Mm.merged, dims = 1:100)
Mm.merged <- FindClusters(object = Mm.merged, resolution = 1.1)
Mm.merged <- RunUMAP(Mm.merged, dims = 1:100)
saveRDS(Mm.merged, file = "Mm_umap_subset_rep.rds")

```


```{r}
saveRDS(Mm_umap, file = "../data_output/surat_objects/Mm_umap_rep.rds")
saveRDS(Mm_umap, file = "Mm_umap_subset_rep.rds")

```
```{r}
Mm_umap <- readRDS("../data_output/surat_objects/Mm_umap_rep.rds")
```

```{r}
Mm.markers <- FindAllMarkers(Mm_umap, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
Mm.markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_logFC)
```

##genes to plot

```{r}
write.csv(Mm.markers, file = "../data_output/surat_objects/Mm_umap_markers_rep.csv")
saveRDS(Mm.markers, file = "../data_output/surat_objects/Mm_umap_markers_rep.rds")
```

```{r}
ifnb.list <- SplitObject(Mm_umap, split.by = "sample")
ifnb.list <- lapply(X = ifnb.list, FUN = SCTransform)
features <- SelectIntegrationFeatures(object.list = ifnb.list, nfeatures = 5000)
ifnb.list <- PrepSCTIntegration(object.list = ifnb.list, anchor.features = features)
```

```{r}
saveRDS(ifnb.list, file = "../data_output/surat_objects/ifnb.list.rds")
saveRDS(features, file = "../data_output/surat_objects/features.rds")



```

```{r}
ifnb.list <- readRDS(file = "../data_output/surat_objects/ifnb.list.rds")
features <- readRDS(file = "../data_output/surat_objects/features.rds")

```

```{r}
immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, normalization.method = "SCT",
    anchor.features = features)
```
```{r}
saveRDS(immune.anchors, file = "../data_output/surat_objects/anchors.rds")

```
```{r}
immune.anchors <- readRDS(file = "../data_output/surat_objects/anchors.rds")
```


```{r}

immune.combined.sct <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")
```

```{r}
saveRDS("combined.sct.rds")
```

#200 is nice
```{r}

immune.combined.sct <- RunPCA(immune.combined.sct, verbose = FALSE, npcs = 300)
ElbowPlot(immune.combined.sct, ndims = 300)
immune.combined.sct <- RunUMAP(immune.combined.sct, reduction = "pca", dims = 1:20)
```

```{r}
DimPlot(immune.combined.sct, reduction = "umap", group.by = "sample")
DimPlot(immune.combined.sct, reduction = "umap", group.by = "geno", label = TRUE, repel = TRUE)
DimPlot(immune.combined.sct, reduction = "umap", label = TRUE)
```
```{r}
saveRDS(immune.combined.sct, file = "../data_output/surat_objects/Mm_umap_integrated_rep.rds")
saveRDS(immune.combined.sct, file = "Mm_umap_integrated_rep.rds")

saveRDS(immune.combined.sct, file = "/Users/ajm108/Desktop/Mm_project/Mm_umap_integrated_subset_rep.rds")

Mm_umap <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_umap_integrated.rds")
Mm_umap <- readRDS("/Users/ajm108/Desktop/Mm_project/Mm_umap_integrated_subset.rds")
```

```{r}
Mm.markers <- FindAllMarkers(immune.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
Mm.markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_logFC)
```

Integration for subset
```{r}
ifnb.list <- SplitObject(Mm.merged, split.by = "sample")

ifnb.list <- lapply(X = ifnb.list, FUN = SCTransform)

features <- SelectIntegrationFeatures(object.list = ifnb.list, nfeatures = 5000)

ifnb.list <- PrepSCTIntegration(object.list = ifnb.list, anchor.features = features)

immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, normalization.method = "SCT",
    anchor.features = features)

immune.combined.sct <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")

immune.combined.sct <- RunPCA(immune.combined.sct, verbose = FALSE, npcs = 300)

immune.combined.sct <- RunUMAP(immune.combined.sct, reduction = "pca", dims = 1:20)

saveRDS(immune.combined.sct, file = "Mm_umap_integrated_subset_rep.rds")

immune.combined.sct <- readRDS("Mm_umap_integrated_subset_rep.rds")
Mm.markers <- FindAllMarkers(immune.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
saveRDS(Mm.markers, file = "Mm_umap_integrated_markers_subset_rep.rds")

```


##genes to plot

```{r}
write.csv(Mm.markers, file = "/Users/ajm108/Desktop/Mm_project/Mm_umap_integrated_markers.csv")
saveRDS(Mm.markers, file = "Mm_umap_integrated_markers_rep.rds")

write.csv(Mm.markers, file = "/Users/ajm108/Desktop/Mm_project/Mm_umap_integrated_markers_subset.csv")
saveRDS(Mm.markers, file = "/Users/ajm108/Desktop/Mm_project/Mm_umap_integrated_markers_subset.rds")
```