---
title: "wnt_pathway_exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(clusterProfiler)
library(org.Mm.eg.db)
```

# Get gene set from gsea
```{r}

mmu_kegg <- clusterProfiler::download_KEGG(species ="mmu")

wnt_pathway <- mmu_kegg$KEGGPATHID2NAME[str_detect(mmu_kegg$KEGGPATHID2NAME$to,"W"), ]

wnt_pathway_entrez_genes <- mmu_kegg$KEGGPATHID2EXTID %>%
  filter(from == wnt_pathway$from )

wnt_pathway_genes <- clusterProfiler::bitr(wnt_pathway_entrez_genes$to,OrgDb = org.Mm.eg.db,fromType = "ENTREZID",toType=c("ENSEMBL","SYMBOL"))


```
# Look at expression in pseudobulk 

Read log norm summed pseudobulk count matrix and filter to get only WNT genes

```{r}
lognormsummed <- readRDS("../data_output/osca_objects/lognormsummed_pseudobulk_de.Rds")

wnt_pathway_genes <- wnt_pathway_genes[wnt_pathway_genes$ENSEMBL %in% rownames(lognormsummed),]

lognormsummed <- lognormsummed[wnt_pathway_genes$ENSEMBL,]



```

Read DE genes from pseudobulk

```{r}
de_results_per_clust_sig <- read_csv("../data_output/surat_objects/deg_pseudobulk_per_cluster_fdr0.05_df.csv")

```

Tidy counts matrix
```{r}
count_mat <- assay(lognormsummed, "logcounts")

colnames(count_mat) <-
paste(lognormsummed$sample,
lognormsummed$geno,
lognormsummed$seurat_clusters,sep   = "_")

counts_df <- as.data.frame(count_mat) %>%
  rownames_to_column(var = "ENSEMBL") %>%
  pivot_longer(cols = -ENSEMBL) %>%
  mutate(sample = str_split(name, "_") %>% map_chr(1),
         geno = str_split(name, "_") %>% map_chr(2),
         seurat_clusters = str_split(name, "_") %>% map_chr(3))

```


Assign gene symbols
```{r}
counts_df <- left_join(counts_df , wnt_pathway_genes)
```

Boxplot
```{r}
counts_df %>%
  ggplot(aes(geno, value, color = sample)) +
  geom_violin() +
  geom_boxplot(alpha = 0.5) +
  theme_light() +
  facet_wrap(~seurat_clusters) +
  scale_color_brewer(type = "div")
  
```
Heatmap
```{r}
library("tidyHeatmap")

counts_df %>%
  filter(seurat_clusters==1) %>%
  heatmap(SYMBOL,name , value,.scale = "row", palette_value = RColorBrewer::brewer.pal(3,"PRGn"),cluster_columns = FALSE) %>%
  add_tile(geno)


```   

```{r}
count_mat_wt1 <- as.data.frame(count_mat) %>%
  dplyr::select(starts_with("WT1"))
count_mat_xp1 <- as.data.frame(count_mat) %>%
  dplyr::select(starts_with("XP1"))
count_mat_wt2 <- as.data.frame(count_mat) %>%
  dplyr::select(starts_with("WT2"))
count_mat_xp2 <- as.data.frame(count_mat) %>%
  dplyr::select(starts_with("XP2"))

diff_df <- cbind(count_mat_wt1 - count_mat_xp1,
count_mat_wt2 - count_mat_xp2)


```


```{r}
diff_df <- diff_df %>%
    rownames_to_column(var = "ENSEMBL") %>%
  pivot_longer(cols = -ENSEMBL) %>%
  mutate(litter = str_split(name, "_") %>% map_chr(1) %>% substr(3,3),
         seurat_clusters = str_split(name, "_") %>% map_chr(3))


diff_df <- left_join(diff_df , wnt_pathway_genes)

```

```{r}

diff_df <- diff_df %>% mutate(name = paste(seurat_clusters, litter))

diff_df_list <- split(diff_df, diff_df$seurat_clusters)
   

imap(diff_df_list, function(.x,.y){
  .x %>%
    heatmap(SYMBOL,name , value,.scale = "none", palette_value = RColorBrewer::brewer.pal(3,"PRGn"), cluster_columns = FALSE, column_title = .y)  

})

```

Are some of the DE in the Wnt pathway?

```{r}
de_results_per_clust_sig[de_results_per_clust_sig$ENSEMBL %in% wnt_pathway_genes$ENSEMBL,]
```

This is a negative regulator of the pathway!
```{r}
de_results_df <- readRDS("../data_output/surat_objects/deg_pseudobulk_per_cluster.Rds")

de_results_df$wnt <- de_results_df$ENSEMBL %in% wnt_pathway_genes$ENSEMBL
```

What is the fold change on average of these genes?
```{r}
de_results_df %>%
  drop_na() %>%
  ggplot(aes(logCPM, logFC)) +
  geom_point(size = 0.5, pch = 1, color = "gray") +
  geom_point(data = drop_na(de_results_df) %>%
               filter(wnt),size = 0.5, pch = 1, color = "blue") +
  scale_color_manual(values = c("TRUE" = "red","FALSE"="darkgray")) +
  facet_wrap(vars(cluster)) +
  theme_light() +
  ylim(c(-2,2)) +
  theme(aspect.ratio = 1) 

  
```

```{r}
de_results_df %>%
  drop_na() %>%
  filter(wnt) %>%
  ggplot(aes(logCPM, logFC, color = -10 * log10(FDR))) +
  geom_point(size = 0.5) +
  scale_color_viridis_c(option = "A") +
  facet_wrap(vars(cluster)) +
  theme_light() +
  ylim(c(-2,2)) +
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 0.5) +
  geom_hline(yintercept = - 0.5) 


  
```

```{r}
de_results_df %>%
  drop_na() %>%
  filter(wnt) %>%
  ggplot(aes(logCPM, logFC, color = FDR < 0.05)) +
  geom_point(size = 0.5) +
  facet_wrap(vars(cluster)) +
  theme_light() +
  theme(aspect.ratio = 1) +
  scale_color_manual(values = c("FALSE" = "gray","TRUE" = "blue"))



```


# Look at expression in single cells 

